Draft -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- Draft

### 2024-09-27 docs/_tex/index.tex
add updates

Coastal areas are vital hotspots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between the land and oceans \citep{unsworth2022}. Seagrass meadows
provide a myriad of ecosystem services to humanity, including carbon
sequestration, oxygen production, protection against sea-level rise and
coastline erosion, and mitigation of eutrophication \citetext{\citealp[
]{unsworth2022}; \citealp{sousa2019blue}}. They serve as vital habitats
for a diverse array of marine and terrestrial species, providing living,
breeding, and feeding grounds \citetext{\citealp[
]{gardner2018}; \citealp[ ]{Zoffoli2022}; \citealp{jankowska2019}}. Due
to the concentration of human activities in coastal zones, seagrass
meadows are directly exposed to and impacted by anthropogenic pressures.
Global regression and fragmentation of seagrass meadows are currently
observed due to climate change, diseases, urbanization, land
reclamation, dredging, competition with alien species, and reduction in
water quality \citetext{\citealp[ ]{nguyen2021}; \citealp[
]{soissons2018}; \citealp[ ]{orth2006}; \citealp[ ]{lin2018}; \citealp[
]{duffy2019}; \citealp[ ]{rasheed2011long}; \citealp[
]{chefaoui2018dramatic}; \citealp{sousa2019blue}}. Both habitat
fragmentation and reduction, in turn, can severely compromise the
effectiveness of ecosystem services provided by seagrass meadows. While
improvements in water quality and hydrodynamics have been recently
reported in Europe, allowing an overall recovery of seagrass ecosystems
at local and European scales, many coastal waters worldwide are still
subjected to strong eutrophication processes \citetext{\citealp[
]{deSantos2019}; \citealp[ ]{Zoffoli2021}; \citealp{sousa2019blue}}.
Coastal eutrophication has been associated to excessive accumulation of
green macroalgae, so-called green tides \citep{devlin2023nutrients}.
Green tides produce shade and suffocation over seagrass individuals,
thus threatening the health of seagrass ecosystems \citep{wang2022}.

@@

Drones (Unmanned Aerial Vehicles -- UAVs) can potentially fill the data
gaps left by satellite RS and \emph{in situ} measurements, due to their
ability to provide spatially-explicit observations at very high spatial
resolutions (pixel size from mm to cm) while capturing data at
multi-spectral resolution \citetext{\citealp[
]{fairley2022drone}; \citealp{oh2017use}}. The versatility of drones
allows for their application across a diverse thematic range , from
coastal zone management \citetext{\citealp[ ]{adade2021}; \citealp[
]{casella2020}; \citealp{angnuureng2022}} to mapping species
distribution \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[ ]{Roman2021}; \citealp[
]{Brunier2022Topographic}; \citealp{sousa2019blue}}. However, when
applied to coastal habitat mapping, most case studies were previously
limited to a single flight, restricting the generalizability of their
application over wider geographical scales \citetext{\citealp[
]{Roman2021}; \citealp[ ]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. The present study
aimed at analyzing the potential of multispectral drone RS to map
intertidal macrophytes over a diverse biogeographical range, with a
particular focus on discriminating Magnoliopsida and Chlorophyceae
(Seagrass and Green Algae, respectively). Ten drone flights were
performed over soft-bottom intertidal areas along the Atlantic
coastlines of two European countries (France and Portugal), covering a
wide range of habitats, from monospecific seagrass meadows to meadows
mixed with green, or red macroalgae. A deep learning algorithm was
trained and validated for macrophyte discrimination, emphasizing
applicability across diverse sites without a loss of prediction
accuracy.

@@

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites were
located in the Gulf of Morbihan (Figure~\ref{fig-map} A), France
(47.5791°N, 2.8018°W). This gulf covers an area of 115 km² and is only
connected to the sea through a 900 m wide channel. A total of 53 small
islands are scattered across the gulf leading to 250 km of shorelines.
Patchy seagrass meadows can be found on many of these islands. One of
the sites within the gulf was on one its islands (Arz) and the other was
located further south on a mainland beach area (Duer). Two other sites
were located in Bourgneuf Bay, France (46.9849°N, 2.1488°W). This is a
340 km² semi-enclosed macrotidal bay, protected from waves by
Noirmoutier Island. Bourgneuf bay hosts a large intertidal seagrass
meadow of about 6 km² \citep{ZOFFOLI2020112020}. Within this meadow, the
sites observed by drones (L'Epine and Barbatre, Figure~\ref{fig-map} B)
contained monospecific beds of \emph{Nanozostera noltei} (dwarf
eelgrass, syn. \emph{Zostera noltei}) with very little mixing with other
macrophytes. Three sites were surveyed in the Ria de Aveiro Coastal
Lagoon in Portugal (40.6887°N, 8.6810°W). The extent of this lagoon is
\textasciitilde83 km² (at low tide) with many narrow channels, large
salt marshes and many mudflats that uncover at low tide
\citep{sousa2017blue}. It is connected to the open sea through a single
channel, with a tidal lag between the North and the South of the lagoon.
The southernmost site (Gafanha) is a mudflat located in the Mira channel
(one of the four main channels of the lagoon) whereas the two other
sites (Mataducos and Marinha Lanzarote) were situated in the middle of
the lagoon and only accessible by boat (Figure~\ref{fig-map} C). These
Portuguese sites are characterized by a more diverse intertidal
vegetation, where patches of seagrass intermingle with red, brown, and
green macroalgae.

@@

Before each flight, targets used as ground control points were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). Ground control points were used to
correct georeferencing imprecision of orthomosaics with an horizontal
and vertical accuracy of 10cm. A dGPS was also used to georeference
quadrats of 0.25 m² ,which assessed the presence or absence of five key
taxonomic classes of intertidal vegetation : Bacillariophyceae
(unicellular benthic diatoms forming biofilms at the sediment surface
during low tide), Phaeophyceae (brown macroalgae), Magnoliopsida (dwarf
eelgrass), Chlorophyceae (green macroalgae) and Rhodophyceae (red
macroalgae) (Figure~\ref{fig-vegetation}). Pictures of each quadrat were
uploaded online to the Global Biodiversity Information Facility (GBIF)
platform \citep{BedeGbif}, a public open portal to store and share
biodiversity data. Each photograph was also processed to estimate the
percent cover of each type of vegetation using an image processing
software \citep[ImageJ,][]{schneider2012nih}. Hyperspectral reflectance
signatures of each vegetation class were recorded using an ASD FieldSpec
HandHeld 2 spectroradiometer, which acquires reflectance between 325 and
1075 nm, with 1 nm of spectral resolution. Hyperspectral signatures
served dual purposes: they validate the radiometric calibration of drone
data and contribute to error reduction in photo interpretations.

@@

The spectral similarities of the reflectance signatures between
intertidal green macrophytes (Magnoliopsida and Chlorophyceae) make
their discrimination challenging using simple classification algorithms
(Figure~\ref{fig-vegetation} F). To overcome this challenge, a deep
learning classification method was trained, validated, and applied to
each drone flight (Figure~\ref{fig-workflow}).

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model. The training pixels were
categorized into seven different classes, representing the various
habitats encountered at the different study sites: Sediment, Water,
Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and
Rhodophyceae. Only low-altitude flights (Table~\ref{tbl-flights}) were
used for training because their 8 mm spatial resolution allowied to
avoid spectral sub-pixel mixing and to accurately identify vegetation
classes. More than 418,000 pixels at 8 mm resolution from the 3 training
flights were used to train the model (Table~\ref{tbl-validationPX}).
Twenty one variables were used by the model as predictors: the ten raw
spectral bands of the Micasense RedEdge Dual MX multispectral camera
(ranging from 444 nm to 840 nm), the same ten spectral bands
standardized using a min/max transformation (Equation~\ref{eq-std} ;
\citep{Cao2017}) and the Normalized difference vegetation index (NDVI,
Equation~\ref{eq-ndvi}). Standardisation of spectral bands is commonly
used to eliminate the scaling differences between spectra and to limit
the effect of biomass on the spectra shape \citetext{\citealp[
]{Douay2022}; \citealp{Davies2023}}.

@@

A neural network classification model was built using the fastai
workflow \citep{howard2018fastai}. This model is composed by 2 hidden
layers and have a total of 26 054 trainable parameters. Parameters have
been find tune using 12 epoch to minimize the error rate.

@@

The model was applied to all flights at both 12 and 120 m of altitude.
In situ information on georeferenced class type and percent cover
collected during each flight was used to assess the model accuracy.
These images were used to construct a validation dataset indicating the
presence or absence of each class. Additionally to the quadrat-based
validation dataset, polygons of each class were photo interpreted in
order to increase the number of pixels of the validation dataset.A total
of 536,000 pixels were used to validate the Neural Network classifier.
The sites with the lowest and highest number of validation data were
Gafanha Low (17316 pixels) and Marinha Lanzarote (159713 pixels),
respectively.A confusion matrix, along with precision metrics such as
global accuracy, sensitivity, specificity, and Kappa coefficient, was
generated for each sites. All validation matrices were then aggregated
to create an overall matrix .

@@

\subsection{Impact of the vegetation cover on the
prediction}\label{impact-of-the-vegetation-cover-on-the-prediction}

The key aspect of the workflow adopted in the present study is the
mapping of specific areas at two different altitudes (12m and 120m),
resulting in two distinct resolutions for the same area (8mm and 80mm).
The predictions made on the high-resolution flight can be used to
estimate the cover (percentage, \%) of each vegetation class for every
pixel of the lower resolution flight. Consequently, for each pixel of
the high-altitude flights, the cover (\%) of each vegetation class were
obtained, and a kernel density plot was generated. This plot provides a
visual representation of the behavior of the model in different
vegetation percent cover scenarios.

@@

@@

Each drone flight was used to produce a prediction map, as well as a
probability map that indicates the model derived likelihood of the
selected class for every pixel. The low-altitude flight conducted in
Gafanha, Portugal, represents the site with the highest complexity
(Figure~\ref{fig-GafLow}). Among the five vegetation classes on which
the model was trained, four were present on this site, with
Chlorophyceae and Rhodophyceae mixed with the seagrass meadow. There was
also Bacillariophyceae forming biofilms on parts of bare sediment
surface. Although the seagrass bed was solely composed of
\emph{Nanozostera noltei}, various colors of this specie could be
observed from dark green (corresponding to healthy leaves) to
whitish/brown (when leaves were discolored having lost their
pigmentation). Regardless of color, Magnoliopsida pixels were accurately
predicted by the model.

@@

The high-altitude flight over Gafanha covered a total area of
\textasciitilde1 km² (Figure~\ref{fig-GafHigh}). A channel contouring a
small island was masked in the prediction map. Most of the intertidal
area was classified as Magnoliopsida by the model, including seagrass
patches with discolored leaves. Only a few pixels were classified as
Chlorophyceae at this spatial scale. Furthermore, the area that was
classified as Bacillariophyceae in the low-altitude flight remained
mostly classified as such in the high-altitude flight, though some
pixels were classified as Magnoliopsida. Patches of Rhodophyceae were
correctly classified. In the northern part of the site and near the land
eadges, patches of the schorre angiosperm \emph{Sporobolus maritimus}
(syn. \emph{Spartina maritima)} were misclassified, either as
Magnoliopsida or as Phaeophyceae.

@@

Amoung the high altitude flights, the one acquired over the inner part
of Ria de Aveiro coastal lagoon covered the largest area with
approximately 1.5 km² (Figure~\ref{fig-Boat}). This site was dominated
by seagrass and red macroalgae. The classification provided consistent
results, with a patchy seagrass meadow mixed with red macroalgae on the
eastern part of the site. As shown in the zoom (Figure~\ref{fig-Boat}),
the edges of the meadow visually appears to be colonised by green
macroalgae (\emph{Ulva sp.}), which the model agreed with.

@@

The flight over L'Epine in Noirmoutier Island, France
(Figure~\ref{fig-Dike}) was conducted near a dike, which crossed the
northern part of the site from west to east. Alongside this dike, Fucale
brown algae (\emph{Fucus spp.}, \emph{Ascophyllum nodosum}) were
attached to sparse rocks and stranded green algae (\emph{Ulva spp.})
could be observed. Despite the high mixture between Chlorophyceae and
Magnoliopsida these two classes were correctly discriminated by the
classifier.

@@

\subsection{Validation}\label{validation-1}

\subsubsection{Reflectance comparison between the two different
altitudes}\label{reflectance-comparison-between-the-two-different-altitudes}

In this study, a key innovation lies in the utilization of drone flights
at two different altitudes (12 m and 120 m) for constructing the neural
network model. Overall there was a good agreement between the two
altitudes (Figure~\ref{fig-CompareRef}).

There was a slight underestimation for raw reflectance values in the
high-altitude flight, particularly for higher reflectance values
(Figure~\ref{fig-CompareRef} A). Since both flights were conducted over
vegetation areas, the highest reflectance values correspond to the
infrared part of the spectrum. This difference is not present when
reflectance values have been standardized (Equation~\ref{eq-std} ;
Figure~\ref{fig-CompareRef} B).

\subsubsection{Neural network classification
validation}\label{neural-network-classification-validation}

Model global accuracy was 94.26\% with a Kappa coefficient of 0.92
(Figure~\ref{fig-Validation}).

@@

Using the very high resolution low altitude flight (8 mm pixels), we
determined the minimal percent cover required to correctly classify a
given class within the corresponding high altitude flight
(Figure~\ref{fig-upscaling}).

@@

When the vegetation cover of a given class was 100 \%, coarser
high-flight pixels were well classified for all the classes except for
Bare Sediment, which was only well classified 80\% of the time. This
phenomenon may be attributed to the time gap between the two flights,
allowing for microphytobenthos migration to the surface during low tide,
consequently altering the model's classification from bare sediment to
Bacillariophyceae. A percent cover of at least 80\% was sufficient to
have all the pixels of high altitude flights correctly classified, with
the exception of Magnoliopsida that required a higher cover
(\textgreater90 \%) to be well classified. Concerning the probability of
each class, the highest percent cover was needed to confidently predict
Bacillariophyceae. To predict Chlorophyceae with a model likelihood of
0.85, a cover of 93 \% was needed, 90 \% for Magnoliopsida, 92 \% for
Rhodophyceae and 97 \% for Bacillariophyceae.

@@

The primary objective of this study was to develop a method for the
accurate classification of macrophytes on intertidal mudflats and
sandflats, specifically focusing on distinguishing between Chlorophyceae
(green macroalgae) and marine Magnoliopsida (seagrasses) using
multispectral drone data. The ability to differentiate between various
types of vegetation plays a critical role in ecological monitoring and
coastal management \citep{WFD2000}. By distinguishing between seagrasses
and macroalgae, our approach facilitates targeted conservation
strategies, enabling more effective preservation and restoration efforts
in coastal ecosystems. The discrimination of seagrasses from green
macroalgae presents two main challenges \citetext{\citealp[
]{oiry2021using}; \citealp[
]{bannari2022}; \citealp{veettil2020opportunities}}. First these two
macrophytes have a similar pigment composition: chlorophyll-a (common to
all vegetation types), chlorophyll-b (an additional photosynthetic
pigment), and accessory carotenoids such as zeaxanthin, lutein and
neoxanthin (Figure~\ref{fig-Pigm}). Second, seagrass and green
macroalgae frequently co-occur in intertidal areas, and can intermingle
within a remote sensing pixel if the spatial resolution is too low.
Here, the issue of intra-pixel mixing was resolved thanks to the very
high spatial resolution of the drone, and the risk of spectral confusion
was avoided thanks to the efficiency of the multispectral classifier.
Our results confirm a recent study based on \emph{in situ} radiometry,
where it was demonstrated that a sensor with at least eight spectral
bands ranging from 500 to 850 nm and including a green band at 530 nm
and a RedEdge band at 730 nm was crucial to accurately discriminate
green macroalgae from seagrasses \citep{Davies2023}.

@@

Meeting these two criteria, the Micasense RedEdge-MX DUAL camera used in
this study, enabled the classifier to achieve 97\% of accuracy between
these two classes (Figure~\ref{fig-ValidationGreen}). Even if the
pigment composition of green macrophytes is similar, differences in the
spectral shape can be observed, with green algae having a higher
reflectance peak at 560 nm as well as a higher NIR plateau than seagrass
(Figure~\ref{fig-vegetation}). Such differences were previously
attributed to differences in pigments concentration and/or ratios
\citep{bargain2013seasonal}, cellular structure as well as in the
orientation of the plant over the sediment surface \citetext{\citealp[
]{beach1997vivo}; \citealp[
]{kirk1994light}; \citealp{hedley2018influence}}.

The variable importance analysis (Figure~\ref{fig-VIP}) identified that
the band at 531 nm was the most important for accurately identifying
Chlorophyceae. In fact, at this wavelength, Chlorophyceae exhibited the
highest reflectance among all other classes, highlighting the difference
in carotenoids to chlorophyll-a ratios between seagrasses and green
macroalgae \citep{repolho2017seagrass}.

>>>>

Coastal areas are vital hotspots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between the land and oceans \citep{unsworth2022}. Seagrass meadows
provide a myriad of ecosystem services, including carbon sequestration,
oxygen production, protection against sea-level rise and coastline
erosion, and mitigation of eutrophication \citetext{\citealp[
]{unsworth2022}; \citealp{sousa2019blue}}. They serve as vital habitats
for a diverse array of marine and terrestrial species, providing living,
breeding, and feeding grounds \citetext{\citealp[
]{gardner2018}; \citealp[ ]{Zoffoli2022}; \citealp{jankowska2019}}. Due
to the concentration of human activities in coastal zones, seagrass
meadows are directly exposed to and impacted by anthropogenic pressures.
Global regression and fragmentation of seagrass meadows are currently
observed due to climate change, diseases, urbanization, land
reclamation, dredging, competition with alien species, and reduction in
water quality \citetext{\citealp[ ]{nguyen2021}; \citealp[
]{soissons2018}; \citealp[ ]{orth2006}; \citealp[ ]{lin2018}; \citealp[
]{duffy2019}; \citealp[ ]{rasheed2011long}; \citealp[
]{chefaoui2018dramatic}; \citealp{sousa2019blue}}. Both habitat
fragmentation and reduction, in turn, can severely compromise the
effectiveness of ecosystem services provided by seagrass meadows. While
improvements in water quality and hydrodynamics have been recently
reported in Europe, allowing an overall recovery of seagrass ecosystems
at local and European scales, many coastal waters worldwide are still
subjected to strong eutrophication processes \citetext{\citealp[
]{deSantos2019}; \citealp[ ]{Zoffoli2021}; \citealp{sousa2019blue}}.
Coastal eutrophication has been associated to excessive accumulation of
green macroalgae, so-called green tides \citep{devlin2023nutrients}.
Green tides produce shade and suffocation over seagrass individuals,
thus threatening the health of seagrass ecosystems \citep{wang2022}.

@@

Drones (Unmanned Aerial Vehicles -- UAVs) can potentially fill the data
gaps left by satellite RS and \emph{in situ} measurements, due to their
ability to provide spatially-explicit observations at very high spatial
resolutions (pixel size from mm to cm) while capturing data at
multi-spectral resolution \citetext{\citealp[
]{fairley2022drone}; \citealp{oh2017use}}. The versatility of drones
allows for their application across a diverse thematic range , from
coastal zone management \citetext{\citealp[ ]{adade2021}; \citealp[
]{casella2020}; \citealp{angnuureng2022}} to mapping species
distribution \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[ ]{Roman2021}; \citealp[
]{Brunier2022Topographic}; \citealp{sousa2019blue}}. However, when
applied to coastal habitat mapping, most case studies were previously
limited to a single flight, restricting the generalizability of their
application over wider geographical scales \citetext{\citealp[
]{Roman2021}; \citealp[ ]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. The present study
aimed at analyzing the potential of multispectral drone RS to map
intertidal macrophytes over a diverse biogeographical range, with a
particular focus on discriminating Magnoliopsida and Chlorophyceae
(Seagrass and Green Algae, respectively). Ten drone flights were
performed over soft-bottom intertidal areas along the Atlantic
coastlines of two European countries (France and Portugal), covering a
wide range of habitats, from monospecific seagrass meadows to meadows
mixed with green, or red macroalgae. A deep learning algorithm was
trained and validated for macrophyte discrimination, emphasizing
applicability across diverse sites without a loss of prediction
accuracy. In addition, the results obtained from drone-based
multispectral remote sensing will be spatially degraded to satellite
resolutions, and the effect of this resolution on the model's
predictions will be investigated, providing insights for coastal habitat
mapping using satellites.

@@

Seven study sites distributed between France and Portugal were selected
for their extensive intertidal seagrass beds. Two sites were located in
the Gulf of Morbihan (Figure~\ref{fig-map} A), France (47.5791°N,
2.8018°W). This gulf covers an area of 115 km² and is only connected to
the sea through a 900 m wide channel. A total of 53 small islands are
scattered across the gulf leading to 250 km of shorelines. Patchy
seagrass meadows can be found on many of these islands. One of the sites
within the gulf was on one its islands (Arz) and the other was located
further south on a mainland beach area (Duer). The Gulf of Morbihan is a
Natura 2000 site and a Regional Protected Area due to its rich
biodiversity, including its seagrass meadows, and is also classified as
a RAMSAR site, which highlights its significance as a wetland of
international importance. Two other sites were located in Bourgneuf Bay,
France (46.9849°N, 2.1488°W). This is a 340 km² semi-enclosed macrotidal
bay, protected from waves by Noirmoutier Island. Bourgneuf bay hosts a
large intertidal seagrass meadow of about 6 km²
\citep{ZOFFOLI2020112020}. Within this meadow, the sites observed by
drones (L'Epine and Barbatre, Figure~\ref{fig-map} B) contained
monospecific beds of \emph{Nanozostera noltei} (dwarf eelgrass, syn.
\emph{Zostera noltei}) with very little mixing with other macrophytes.
Bourgneuf Bay is also part of the Natura 2000 network and serves as a
RAMSAR site due to its critical habitat for migratory bird species and
its extensive seagrass meadows. Three sites were surveyed in the Ria de
Aveiro Coastal Lagoon in Portugal (40.6887°N, 8.6810°W). The extent of
this lagoon is \textasciitilde83 km² (at low tide) with many narrow
channels, large salt marshes and many mudflats that uncover at low tide
\citep{sousa2017blue}. It is connected to the open sea through a single
channel, with a tidal lag between the North and the South of the lagoon.
The southernmost site (Gafanha) is a mudflat located in the Mira channel
(one of the four main channels of the lagoon) whereas the two other
sites (Mataducos and Marinha Lanzarote) were situated in the middle of
the lagoon and only accessible by boat (Figure~\ref{fig-map} C). These
Portuguese sites are characterized by a more diverse intertidal
vegetation, where patches of seagrass intermingle with red, brown, and
green macroalgae. The Aveiro Lagoon, like the other study areas, is a
Natura 2000 site and a RAMSAR wetland, recognized for its rich mosaic of
habitats and importance for biodiversity, including migratory bird
species and intertidal vegetation.

@@

Before each flight, targets used as ground control points were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). Ground control points were used to
correct georeferencing imprecision of orthomosaics with an horizontal
and vertical accuracy of 10cm. A dGPS was also used to georeference
quadrats of 0.25 m² ,which assessed the presence or absence of five key
taxonomic classes of intertidal vegetation : Bacillariophyceae
(unicellular benthic diatoms forming biofilms at the sediment surface
during low tide), Phaeophyceae (brown macroalgae), Magnoliopsida (dwarf
eelgrass), Chlorophyceae (green macroalgae) and Rhodophyceae (red
macroalgae) (Figure~\ref{fig-vegetation}). Only homogeneous vegetation
patches extending over several meters were selected as ground control
points. Pictures of each quadrat were uploaded online to the Global
Biodiversity Information Facility (GBIF) platform \citep{BedeGbif}, a
public open portal to store and share biodiversity data. Each photograph
was also processed to estimate the percent cover of each type of
vegetation using an image processing software
\citep[ImageJ,][]{schneider2012nih}. Hyperspectral reflectance
signatures of each vegetation class were recorded using an ASD FieldSpec
HandHeld 2 spectroradiometer, which acquires reflectance between 325 and
1075 nm, with 1 nm of spectral resolution. Hyperspectral signatures
served dual purposes: they validate the radiometric calibration of drone
data and contribute to error reduction in photo interpretations.

@@

The spectral similarities of the reflectance signatures at the spectral
resolution of the micasense senor between intertidal green macrophytes
(Magnoliopsida and Chlorophyceae) make their discrimination challenging
using simple classification algorithms (Figure~\ref{fig-vegetation} F).
To overcome this challenge, a deep learning classification method was
trained, validated, and applied to each drone flight
(Figure~\ref{fig-workflow}).

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model. The training pixels were
categorized into seven different classes, representing the various
habitats encountered at the different study sites: Sediment, Water,
Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and
Rhodophyceae. Only low-altitude flights (Table~\ref{tbl-flights}) were
used for training because their 8 mm spatial resolution allowed to avoid
spectral sub-pixel mixing and to accurately identify vegetation classes.
More than 418,000 pixels at 8 mm resolution from the 3 training flights
were used to train the model (Table~\ref{tbl-validationPX}). Twenty one
variables were used by the model as predictors: the ten raw spectral
bands of the Micasense RedEdge Dual MX multispectral camera (ranging
from 444 nm to 840 nm), the same ten spectral bands standardized using a
min/max transformation (Equation~\ref{eq-std} ; \citep{Cao2017}) and the
Normalized difference vegetation index (NDVI, Equation~\ref{eq-ndvi}).
Standardisation of spectral bands is commonly used to eliminate the
scaling differences between spectra and to limit the effect of biomass
on the spectra shape \citetext{\citealp[
]{Douay2022}; \citealp{Davies2023}}.

@@

A neural network classification model was built using the fastai
workflow \citep{howard2018fastai}. This model is composed by 2 hidden
layers and have a total of 26 054 trainable parameters. Parameters have
been fine tuned using 12 epoch to minimize the error rate.

@@

The model was applied to all flights at both 12 and 120 m of altitude.
\emph{In situ} information on georeferenced class type and percent
cover, acquired at over homogeneous vegetation patches at the same time
as drone flights was used to assess the model accuracy. These images
were used to construct a validation dataset indicating the presence or
absence of each class. Additionally to the quadrat-based validation
dataset, polygons of each class were photo interpreted in order to
increase the number of pixels of the validation dataset.A total of
536,000 pixels were used to validate the Neural Network classifier. The
sites with the lowest and highest number of validation data were Gafanha
Low (17316 pixels) and Marinha Lanzarote (159713 pixels), respectively.
A confusion matrix, along with precision metrics such as global
accuracy, sensitivity, specificity, and Kappa coefficient, was generated
for each sites. All validation matrices were then aggregated to create
an overall matrix .

@@

\subsection{Impact of mixed vegetation cover on the
prediction}\label{impact-of-mixed-vegetation-cover-on-the-prediction}

The key aspect of the workflow adopted in the present study is the
mapping at two different altitudes (12 m and 120 m), resulting in two
distinct resolutions for the same area (8 mm and 80 mm). The
high-resolution flight was used to estimate the sub-pixel composition
for each pixel of the lower-resolution flight. Consequently, within each
pixel of the high-altitude flights, the contribution of each vegetation
class (\% cover) was obtained, and a kernel density plot was generated.
This plot provided a visual representation of the model's behavior in
mixed vegetation scenarios. It helps to understand the minimum
vegetation cover of a given class within a pixel necessary for the model
to confidently predict that class.

@@

\subsection{Reflectance comparison between the two different
altitudes}\label{reflectance-comparison-between-the-two-different-altitudes}

In this study, drone flights were conducted at two different altitudes
(12 m and 120 m) to construct the neural network model. At the sites
where the flights at both altitudes overlapped, reflectance was
compared. Overall there was a good agreement between the two altitudes
(RMSE: 0.027 ; Figure~\ref{fig-CompareRef}).

There was a slight underestimation for raw reflectance values in the
high-altitude flight, particularly for higher reflectance values
(Figure~\ref{fig-CompareRef} A). Since both flights were conducted over
vegetation areas, the highest reflectance values correspond to the
infrared part of the spectrum. This difference was not present when
reflectance values have been standardized (Equation~\ref{eq-std} ;
Figure~\ref{fig-CompareRef} B).

@@

Each drone flight was used to produce a prediction map, as well as a
probability map that indicate the model derived likelihood of the
selected class for every pixel. The low-altitude flight conducted in
Gafanha, Portugal, represented the site with the highest complexity
(Figure~\ref{fig-GafLow}). Among the five vegetation classes on which
the model was trained, four were present on this site, with green an red
macroalgae mixed with a seagrass meadow. There was also diatoms forming
biofilms on bare sediment surface. Although the seagrass was solely
composed of a single specie, \emph{Nanozostera noltei}, various colors
of this specie could be observed from dark green (corresponding to
healthy leaves) to brown (when leaves are senescent or have an altered
pigment composition). Regardless of the variation of color, the class
Magnoliopsida was accurately predicted by the model (F1 score of 0.96 at
that site).

@@

The high-altitude flight over Gafanha covered a total area of
\textasciitilde1 km² (Figure~\ref{fig-GafHigh}). A channel contouring a
small island was masked in the prediction map. Most of vegetation area
was classified as seagrass by the model, including patches with
discolored leaves. Only a few pixels were classified as green algae (F1
score of 0.55). Patches of Rhodophyceae were correctly classified (F1
score of 0.85). In the northern part of the site and near the land
eadges, patches of the schorre angiosperm \emph{Sporobolus maritimus}
(syn. \emph{Spartina maritima)} were misclassified, either as seagrass
or as brown algae (F1 score of 0.77 and 0.71, respectively).

@@

Amoung the high altitude flights, the one acquired over the inner part
of Ria de Aveiro coastal lagoon covered the largest area with
approximately 1.5 km² (Figure~\ref{fig-Boat}). This site was dominated
by seagrass and red macroalgae. The classification provided consistent
results, with a patchy seagrass meadow mixed with red macroalgae on the
eastern part of the site. As shown in the zoom (Figure~\ref{fig-Boat}),
the edges of the meadow were mixed with green macroalgae (\emph{Ulva
sp.}), which the model agreed with (F1 score of 0.89 for green algae,
0.97 for seagrass and 0.98 for red algae).

@@

The flight over L'Epine in Noirmoutier Island, France
(Figure~\ref{fig-Dike} A) was conducted near a dike, which crossed the
northern part of the site from West to East. Alongside this dike, Fucale
brown macroalgae (\emph{Fucus spp.}, \emph{Ascophyllum nodosum}) were
attached to sparse rocks and stranded green algae (\emph{Ulva spp.})
could be observed and this was correctly reproduced by the prediction
(Figure~\ref{fig-Dike} B). This site was characterized by a high mixture
between green macroalgae and seagrass but these two classes were
correctly discriminated by the classifier (F1 score of 0.97 and 0.98
respectively).

@@

\subsection{Validation of the model}\label{validation-of-the-model}

With all drone flights combined, the model global accuracy was 94.26\%
with a Kappa coefficient of 0.92 (Figure~\ref{fig-Validation}).

@@

Using the very high resolution low altitude flight (8 mm pixels), we
determined the minimal percent cover required to correctly classify a
given class within the corresponding high altitude flight (8cm pixel
resolution ; Figure~\ref{fig-upscaling}).

@@

A percent cover of at least 80\% was sufficient to have all the pixels
of high altitude flights correctly classified, with the exception of
Magnoliopsida that required a higher cover (\textgreater90 \%) to be
well classified. Concerning the probability of each class, there is a
linear relationship between the percent cover and the confidence of the
model to predict the class. To predict Chlorophyceae with a model
likelihood of 0.85, a cover of 93 \% was needed, 90 \% for
Magnoliopsida, 92 \% for Rhodophyceae and 97 \% for Bacillariophyceae.
When the vegetation cover of a given class was 100 \%, coarser
high-flight pixels were well classified for all the classes except for
Bare Sediment, which was only well classified 80\% of the time. This
phenomenon may be attributed to the time gap between the two flights,
allowing for microphytobenthos migration to the surface during low tide,
consequently altering the model's classification from bare sediment to
Bacillariophyceae.

@@

The primary objective of this study was to develop a method for the
accurate classification of macrophytes at low tide on tidal flats,
specifically focusing on distinguishing between Chlorophyceae (green
macroalgae) and marine Magnoliopsida (seagrasses) using a multispectral
resolution. The discrimination of seagrasses from green macroalgae
presents is spectrally challenging \citetext{\citealp[
]{oiry2021using}; \citealp[
]{bannari2022}; \citealp{veettil2020opportunities}}. These two
macrophytes share a similar pigment composition: chlorophyll-a (common
to all vegetation types), chlorophyll-b (an additional photosynthetic
pigment), and accessory carotenoids such as zeaxanthin, lutein and
neoxanthin (Figure~\ref{fig-Pigm}). Their spectreal responses could be
close, particularly at a multispectral resolution. Seagrass and green
macroalgae frequently co-occur in intertidal areas, and can intermingle
within a remote sensing pixel if the spatial resolution is too low.
Here, the issue of intra-pixel mixing was resolved thanks to the very
high spatial resolution of the drone. In this study the risk of spectral
confusion was avoided with a machine-learning approach exploiting a
neural networks classifier. Our drone flights and a recent study based
on \emph{in situ} radiometry, suggested that a sensor with at least
eight spectral bands ranging from 500 to 850 nm and including a green
band at 530 nm and a RedEdge band at 730 nm was crucial to accurately
discriminate green macroalgae from seagrasses \citep{Davies2023}.

@@

Meeting these two criteria, the Micasense RedEdge-MX DUAL camera used in
this study, enabled the classifier to achieve 97\% of accuracy between
these two classes (Figure~\ref{fig-ValidationGreen}). Even if their
pigment composition is similar, differences in the spectral shape can be
observed, with green algae having a higher reflectance peak at 560 nm as
well as a higher NIR plateau than seagrass
(Figure~\ref{fig-vegetation}). Such differences were previously
attributed to differences in pigments concentration and/or ratios
\citep{bargain2013seasonal}, cellular structure as well as in the
orientation of the plant at the sediment surface \citetext{\citealp[
]{beach1997vivo}; \citealp[
]{kirk1994light}; \citealp{hedley2018influence}}.
